---
title: "Outliers"
subtitle: "Check for outliers"
date: 2023-04-04
date-modified: 2023-04-04
order: 6
author: "Chunzhe ZHANG(Spencer)"
toc: true
number-sections: true
highlight-style: pygments
---

# Check Outliers

## Load Library  {.unnumbered}
```{r load_library, echo=FALSE, message=FALSE, warning=FALSE}
require(snt)
require(qreport)
require(Hmisc)
require(qs)
require(readr)
source("paths.R")
options(prType='html')
hookaddcap()
```

## Load Data {.unnumbered}

```{r load_data, message=FALSE, warning=FALSE}
d <- qread(o_routine)
setDT(d)
t <- read_csv(o_routine_csv)
```

## Outliers

Outliers can be defined as particularly influential observations. Most methods rely on the computation of some distance metric, and the observations greater than a certain threshold are considered outliers. Importantly, outliers detection methods are meant to provide information to consider for the researcher, rather than to be an automatized procedure which mindless application is a substitute for thinking. Always review the outliers before take any actions.

```{r fig_outliers, message=FALSE, warning=FALSE}
#| label: fig-outliers
#| fig-cap: “Check outliers”
#| fig-height: 8
#| fig-width: 12
#| layout-ncol: 1
#| fig-subcap:
#| - “All cause outpatients”
#| - “Malaria deaths, malaria treatment, confirmed cases and tests”
#| - “Malaria admission, all cause admission”
x1 <- t |> plot_outliers(allout_ov5,
                   allout_u5,
                   allout)

ggplotlyr(x1) |>
  plotly::toWebGL()

x2 <- t |> plot_outliers(maldth_ov5,
                   maldth_u5,
                   maldth,
                   maltreat_ov5,
                   maltreat_u5,
                   maltreat,
                   conf,
                   test)

ggplotlyr(x2) |>
  plotly::toWebGL()

x3 <- t |> plot_outliers(maladm,
                   maladm_ov5,
                   maladm_u5,
                   alladm,
                   alladm_ov5,
                   alladm_u5)

ggplotlyr(x3) |>
  plotly::toWebGL()
```

Detailed list for outliers
```{r outliers, echo=FALSE, message=FALSE}
kabl(
  t |> find_outliers(
    alpha = 0.999999999999,
    both_sides = FALSE,
    adm0, adm1, adm2, adm3, hfname, year, month
  ) |> dplyr::select(-id)
)
```
