---
title: "Nigeria 2022 SNT"
editor_options:
  markdown:
    wrap: 72
output:
  html_document:
    number_sections: true
    df_print: paged
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  comment = '', fig.width = 6, fig.height = 6,
  tidy = 'formatR',
  cache = TRUE
)
```

```{r include=FALSE}
devtools::load_all()
library(snt)
library(tidyverse)
library(haven)
library(readxl)
library(skimr)
```

```{r}
snt_country <<- "NGA"
```

# 1. cleaning

## 1.0 meta

### 1.0.1 input

```{r}
path_data <- "/Users/sepmein/Library/CloudStorage/OneDrive-SharedLibraries-WorldHealthOrganization/GMP-SIR\ -\ Country_Analytical_Support/Countries/NGA/WHO_NGA/NGA_2022_SNT/_Submitted_data/"
```

#### 1.0.1.1 Routine

```{r}
path_data_routine <- file.path(path_data, "Routine\ data")
path_data_routine_hf_1420 <- file.path(
    path_data_routine,
    "NGA_hflevel_2014-2020_raw.dta"
)
path_data_routine_hf_21 <- file.path(
    path_data_routine,
    "monthly_hf_2021.xlsx"
)
path_data_raw <- file.path(
  path_data_routine,
  "NGA_routineintdata_hf_month.dta"
)
```

### 1.0.2 output

```{r}
path_output_data <- "/Users/sepmein/Library/CloudStorage/OneDrive-SharedLibraries-WorldHealthOrganization/GMP-SIR\ -\ Country_Analytical_Support/Countries/NGA/WHO_NGA/NGA_2022_SNT/Final\ database\ for\ analysis"
file_output <- file.path(
  path_output_data,
  "hflevel_14-21.csv"
)
file_output_fixed_outlier <- file.path(
  path_output_data,
  "hflevel_14-21-fixed_outlier.csv"
)
outliers_output <- file.path(
  path_output_data,
  "outliers_14-21.csv"
)
fixed_outliers_output <- file.path(
  path_output_data,
  "fixed_outliers_14-21.csv"
)
```

### 1.0.3 variable columns
```{r}
variable_columns <- c(  
    "attendance",
    "opd",
    "fever_u5",
    "fever_a5",
    "tested_rdt_u5",
    "tested_rdt_a5",
    "rdtpos_u5",
    "rdtpos_a5",
    "tested_micro_u5",
    "tested_micro_a5",
    "micropos_u5",
    "micropos_a5",
    "clinical_u5",
    "clinical_a5",
    "conf_u5",
    "conf_a5",
    "severe_u5",
    "severe_a5",
    "conf_act_u5",
    "conf_act_a5",
    "clinical_act_u5",
    "clinical_act_a5",
    "conf_othantimal_u5",
    "conf_othantimal_a5",
    "ipd",
    "mal_death_u5",
    "anc_total",
    "anc_b20wks",
    "anc_a20wks",
    "ipt1",
    "ipt2",
    "preg_llin",
    "fi_1yr_fixed",
    "fi_1yr_outreach")
```

## 1.1 merging before 2021 and the 2021 data

### 1.1.1 read health facility monthly data from 2014 - 2021

```{r}
d_hf_14_20 <- read_dta(path_data_routine_hf_1420)
```

#### Get colnames and skim the data

```{r}
colnames(d_hf_14_20)
skimr::skim(d_hf_14_20)
```

#### Show first rows

```{r}
head(d_hf_14_20)
```

### 1.1.2 read health facility monthly data of the year 2021

```{r echo = FALSE}
d_hf_21 <- read_xlsx(path_data_routine_hf_21)
```

#### Get colnames and skim the data

```{r}
colnames(d_hf_21)
skimr::skim(d_hf_21)
```

#### Show first N rows

```{r}
head(d_hf_21)
```

### 1.1.3 Data Ibrahim submitted on 11/06

```{r}
d_hf_21_raw <- read_dta(path_data_raw)
```

```{r}
colnames(d_hf_21_raw)
```

Merge with d_hf_21

```{r}
d_hf_21 |> left_join(d_hf_21_raw)
```

### 1.1.3 Reconstruct hf 2021

#### Rename the columns

```{r}
d_hf_21 <- d_hf_21 |>
  select(
    -fever_u5,
    -fever_a5,
    -tested_rdt_u5,
    -tested_rdt_a5,
    -rdtpos_u5,
    -rdtpos_a5,
    -conf_a5,
    -clinical_u5,
    -clinical_a5,
    -severe_u5,
    -severe_a5,
  ) |>
  rename(
    opd = allout,
    fever_u5  = susp_u5,
    fever_a5  = susp_ov5,
    tested_rdt_u5  = test_rdt_u5,
    tested_rdt_a5  = test_rdt_ov5,
    rdtpos_u5  = conf_rdt_u5,
    rdtpos_a5  = conf_rdt_ov5,
    tested_micro_a5  = test_mic_ov5,
    tested_micro_u5  = test_mic_u5,
    micropos_u5  = conf_mic_u5,
    micropos_a5  = conf_mic_ov5,
    #  clinical_preg  = pres_preg,
    clinical_u5  = pres_u5,
    clinical_a5  = pres_ov5,
    conf_a5  = conf_ov5,
    severe_a5  = maladm_ov5,
    severe_u5  = maladm_u5,
    ipd  = alladm,
    mal_death_u5  = maldth_u5,
    conf_u5 = "conf_u5...61",
    ipt1 = iptp1,
    ipt2 = iptp2,
    ipt3 = iptp3,
    ipt4 = iptp4,
    preg_llin = llins_anc
  )
```

#### Reorder the columns

```{r}
d_hf_21 <- d_hf_21 |>
  select(
    "facilityuid",
    "periodid",
    "periodcode",
    "perioddesc",
    "attendance",
    "opd",
    "fever_u5",
    "fever_a5",
    "tested_rdt_u5",
    "tested_rdt_a5",
    "rdtpos_u5",
    "rdtpos_a5",
    "tested_micro_u5",
    "tested_micro_a5",
    "micropos_u5",
    "micropos_a5",
    #"clinical_preg", # wait for Ibrahim to share data
    #"conf_preg", # wait for Ibrahim to share data
    "clinical_u5",
    "clinical_a5",
    "conf_u5",
    "conf_a5",
    "severe_u5",
    "severe_a5",
    "conf_act_u5",
    "conf_act_a5",
    "clinical_act_u5",
    "clinical_act_a5",
    "conf_othantimal_u5",
    "conf_othantimal_a5",
    "ipd",
    "mal_death_u5",
    #"stockout_act", # no need
    #"stockout_rdt", # no need
    #"stockout_sp", # no need
    #"stockout_llin", # no need
    "anc_total",
    "anc_b20wks",
    "anc_a20wks",
    "ipt1", 
    "ipt2", 
    "preg_llin",
    "fi_1yr_fixed",
    "fi_1yr_outreach",
    #"u5llin", 
    #"mal_ro", 
    #"mal_adr", 
    #stateuid, 
    #lgauid, 
    facility_name,
    "ownership",
    #"facilityname",
    #"longid",
    "year",
    "month",
    "state",
    "lga",
    "ward",
    "periodname"
  )
```

#### Fix the *periodname* column problem

When combining rows of two database: Error in `bind_rows()`: ! Can't
combine `..1$periodname` <date> and `..2$periodname` <character>. fix
that firstly

I found that the date information was already well stored in the
`periodcode` column, so what I am going to do is to **drop**
`periodname` column for both database.

```{r}
d_hf_14_20 <- d_hf_14_20 |> select( -periodname)
d_hf_21 <- d_hf_21 |> select(-periodname)
```

### 1.1.3 Combine Rows for the two dbs

bind rows between the old and new data

```{r}
d_hf_14_21 <- bind_rows(d_hf_14_20, d_hf_21)
```

### 1.1.4 Treat blank as NA

```{r}
d_hf_14_21 <- d_hf_14_21 |> mutate_all(na_if,"")

```

## 1.2 Cleaning

### 1.2.1 state and stateuid

As we can see from the merged data

```{r}
state_db <- d_hf_14_21 |>
  distinct(
    state,
    stateuid
  ) |>
  arrange(state)
state_db
```

We need to clean it a little bit. 1. Get a standard state name and uid
2. replace\_ it with the incorrect names 3. assign correct stateuid to
the according column

```{r}
# get a standard state name and uid

# remove unneccesary lines
state_db <- state_db |> 
  filter(!is.na(stateuid)) |> 
  filter(!nchar(stateuid) == 0)
  # use snt::routine_replace function to replace_ the state name
state_db <- snt::import_replace(
  state_db,
  snt::import_routine_adm1_replace,
  snt_country = snt_country,
  column = "state"
) |>
  distinct()
state_db
```

Now only have 37 rows.

Check the current \# of states in the db.

```{r}
d_hf_14_21 |> select(state) |> distinct() |> arrange(state)
```

Next, replace\_ all the state names in the database,

```{r}
d_hf_14_21 <- snt::import_replace(d_hf_14_21,
                                  snt::import_routine_adm1_replace,
                                  "NGA",
                                  "state") 

d_hf_14_21 |>
  select(state) |>
  distinct() |>
  arrange()
```

Remove the stateuid column then merge back the stata_db.

```{r}
d_hf_14_21 <- d_hf_14_21 |> 
  select(-stateuid) |>
  left_join(state_db)

d_hf_14_21 |> skim(stateuid)
```

Now 100% complete for stateuid. Fixed.

### 1.2.2 LGA names

#### LGA

Check the \# LGAs firstly.

```{r}
d_hf_14_21 |> select(lga) |> distinct() |> arrange(lga)
```

There are 1560 different LGAs for the original data.

The naming system in NGA for LGAs is like: "ak Mbo Local Government
Area", this should be changed to "Mbo"

```{r}
d_hf_14_21 <- d_hf_14_21 |>
  mutate_at(c("lga"), ~str_remove(., "^([a-z]{2} )")) |>
  mutate_at(c("lga"), ~str_remove(., " Local Government Area")) 
```

Check if it's better now:

```{r}
d_hf_14_21 |> distinct( lga ) |> arrange(lga)
```

I have created a database inside `snt` package for renaming, along with
a very useful function called `snt::import_replace` for easy replacing.

Standard `LGA/adm1` replacement from `snt` package looks like:

```{r}
nga_import_routine_adm2_replace <- snt::import_routine_adm2_replace |>
  filter(country == snt_country) |> arrange(from)
nga_import_routine_adm2_replace
```

Next, use the `snt::import_routine_adm2_replace` database to replace the
LGA column.

```{r}
# get routine rename database from snt database
d_hf_14_21 <- snt::import_replace(
  d_hf_14_21,
  snt::import_routine_adm2_replace,
  snt_country = snt_country,
  column = "lga"
)
```

Check if it's better now:

```{r}
d_hf_14_21 |> distinct( lga ) |> arrange(lga)
```

Next step, we need to rename the LGAs to align with those in shapefile.
The problem for NGA is sometimes, different adm1(state) has same LGA
name. E.g. both *Kogi* and *Plateau* has a LGA called *Bassa*. They
should be renamed with a number. country \| adm1 \| from \| to NGA \|
Kogi \| Bassa \| Bassa1\
NGA \| Plateau \| Bassa \| Bassa2

I have also recreated a db called
`snt::import_routine_adm2_replace_by_adm1`

replace LGAs to align the names in shapefile

```{r}
snt::import_routine_adm2_replace_by_adm1
```

Use `snt::import_replace` fn to replace those names.

```{r}
d_hf_14_21 <- snt::import_replace(
  d_hf_14_21,
  snt::import_routine_adm2_replace_by_adm1,
  snt_country = snt_country,
  column = "lga",
  by = c("state" = "adm1")
)
```

We can see the \# of lgas increased.

```{r}
d_hf_14_21 |> distinct(lga) |> arrange(lga)
```

#### TODO LGA UID

Send to #Ibrahim ask for why.

As 21 data do not have LGAs, I need to fix the problem. Skim over LGA and lgauid column firstly.

```{r}
d_hf_14_21 |> skim(lga, lgauid)
```

Number of missing `lgauid` is 1677208

Print out \# of unique `lgauid` and `lga`.

```{r}
d_hf_14_21 |> distinct(lga, lgauid) |> arrange(lga)
```

As lots of lgauid are missing, filter those blank lines.

```{r}
lga_db <- d_hf_14_21 |> distinct(lga, lgauid) |> 
  arrange(lga) |>
  filter(!is.na(lgauid)) |> 
  filter(!nchar(lgauid) == 0)
lga_db
```

Now 774 rows.

Remove the `lgauid` column in `d_hf_14_21` db, and merge it back with
`lga_db`

```{r}
d_hf_14_21 <- d_hf_14_21 |> select(-lgauid) |> left_join(lga_db) 
d_hf_14_21 |>
  distinct(lga, lgauid) |> 
  arrange(lga) 
```

No missing LGAs, but found a few blank LGA

Print out those lines.

```{r}
d_hf_14_21 |> select(-lgauid) |> left_join(lga_db) |>
  filter(is.na(lga))
```

Try to find where they belongs to Search for 14-20 db

```{r}
facility_db <- d_hf_14_21 |>
  select(facilityuid,
         lga,
         lgauid) |>
  distinct()
facility_db
```

```{r}
d_hf_14_21 |>  
  select(
         -lga,
         -lgauid) |>
  left_join(facility_db) |>
  select(facilityuid,
         lga,
         lgauid) |>
  skim()
```

Still 47676 missing, failed

### 1.2.3 HF

Number of `facility_name` and `facilityuid` is:

```{r}
d_hf_14_21 |> 
  select(facility_name, facilityuid) |>
  distinct()
```

There are some facilities, which is actually the same one but has different names, merging those naming.

```{r}
unique_facilities <- d_hf_14_21 |>
  select(facility_name, facilityuid) |>
  distinct(facilityuid, .keep_all = TRUE)

d_hf_14_21 <- d_hf_14_21 |> 
  select(-facility_name) |>
  left_join(unique_facilities)

d_hf_14_21 |>
  select(facility_name, facilityuid) |>
  distinct()
```

#### The problem of OWNERSHIP

Health Facilities should be either Private or Public, which means adding ownership would not add number of unique health facilities.

While, in reality, running this code gives us 99231 rows of records, instead, hf should be around 50000~.

TODO: ask @Bea whether this info is correct.

```{r}
d_hf_14_21 |>
  select(   
    state,
    stateuid,
    lga,
    lgauid,
    facility_name,
    facilityuid,
    ownership) |>
  distinct()
```

#### Aggregation

Aggregate by health facility, year, month
Dropped the `ward` and `ownership` here.

```{r}
d_hf_14_21 <- d_hf_14_21 |>
  select(
    state,
    stateuid,
    lga,
    lgauid,
    facility_name,
    facilityuid,
    -ownership,
    -facilityname,
    -longid,
    year,
    month,
    -ward,
    #-periodname,
    -periodid,
    -periodcode,
    -perioddesc,
    attendance,
    opd,
    fever_u5,
    fever_a5,
    tested_rdt_u5,
    tested_rdt_a5,
    rdtpos_u5,
    rdtpos_a5,
    tested_micro_u5,
    tested_micro_a5,
    micropos_u5,
    micropos_a5,
    -clinical_preg,
    -conf_preg,
    clinical_u5,
    clinical_a5,
    conf_u5,
    conf_a5,
    severe_u5,
    severe_a5,
    conf_act_u5,
    conf_act_a5,
    clinical_act_u5,
    clinical_act_a5,
    conf_othantimal_u5,
    conf_othantimal_a5,
    ipd,
    mal_death_u5,
    -stockout_act,
    -stockout_rdt,
    -stockout_sp,
    -stockout_llin,
    anc_total,
    anc_b20wks,
    anc_a20wks,
    ipt1,
    ipt2,
    preg_llin,
    fi_1yr_fixed,
    fi_1yr_outreach,
    -u5llin,
    -mal_ro,
    -mal_adr,
  ) |>
  group_by(state,
           stateuid,
           lga,
           lgauid,
           facility_name,
           facilityuid,
           year,
           month) |>
  summarise_all( ~ sum(.x, na.rm = FALSE))
```

# 2.0 Export

Export to the right position.

```{r}
d_hf_14_21 |> write_csv(file_output)
```

# 3.0 Quality and Consistency checks

```{r}
d_hf_14_21 <- read_csv(file_output)
```

### 3.1 Report rate - Plot by HF & Date

Generate yearmon for plotting

```{r}
d_hf_14_21 <- d_hf_14_21 |> 
  mutate(
    yearmon = str_c(year, str_pad(month, 2, pad = "0"))
  )
```

```{r}
report_status_by_hf_and_date <- d_hf_14_21 |>
  group_by(state,
           stateuid,
           lga,
           lgauid,
           facility_name,
           facilityuid,
           year,
           month) |>
  summarise(sum = sum(
    c_across(attendance:fi_1yr_outreach), 
    na.rm = TRUE)) |>
  mutate(reported = if_else(sum > 0, "Y", "N"))

report_status_by_hf_and_date |>
  mutate(yearmon = str_c(year, str_pad(month, 2, pad = "0"))) |>
  ggplot(aes(x = yearmon, y = facility_name, fill = reported)) +
  geom_tile() +
  scale_x_discrete(guide = guide_axis(check.overlap = TRUE)) +
  labs(title = "Report Status By Health Facilitis and Date",
       y = "Health Facilities",
       x = "Date") +
  theme(axis.text.y = element_blank()) +
  scale_fill_manual(values = c("N" = "white", "Y" = "blue"))
```

### 3.2 Report HFs by Adm1

```{r}
report_rate_by_adm1 <- d_hf_14_21 |>
  group_by(state,
           stateuid,
           lga,
           lgauid,
           facility_name,
           facilityuid,
           year,
           month) |>
  summarise(sum = sum(c_across(attendance:fi_1yr_outreach),
                      na.rm = TRUE)) |>
  mutate(reported = if_else(sum >= 0, 1, 0)) |>
  group_by(state) |>
  summarise(report_rate = 1 - sum(reported) / n()) |>
  arrange(desc(report_rate))

ggplot(report_rate_by_adm1) +
  geom_col(aes(y = fct_reorder(state, report_rate), x = report_rate), orientation =
             "y") +
  labs(title = "Percentage of Never Reported HFs",
       y = "Adm1",
       x = "Percentage")
```

### 3.2 Outliers

```{r}
outliers <- find_outlier(
  d_hf_14_21,
  columns = variable_columns,
  alpha = 0.99999,
  select_rows = c(
          "state",
           "stateuid",
           "lga",
           "lgauid",
           "facility_name",
           "facilityuid",
           "year",
           "month"
           )
)
outliers |> write_csv(outliers_output)
```

```{r}
d_hf_14_21 |> skim(
    "attendance",
    "opd",
    "fever_u5",
    "fever_a5",
    "tested_rdt_u5",
    "tested_rdt_a5",
    "rdtpos_u5",
    "rdtpos_a5",
    "tested_micro_u5",
    "tested_micro_a5",
    "micropos_u5",
    "micropos_a5",
    "clinical_u5",
    "clinical_a5",
    "conf_u5",
    "conf_a5",
    "severe_u5",
    "severe_a5",
    "conf_act_u5",
    "conf_act_a5",
    "clinical_act_u5",
    "clinical_act_a5",
    "conf_othantimal_u5",
    "conf_othantimal_a5",
    "ipd",
    "mal_death_u5",
    "anc_total",
    "anc_b20wks",
    "anc_a20wks",
    "ipt1",
    "ipt2",
    "preg_llin",
    "fi_1yr_fixed",
    "fi_1yr_outreach"
)
```

Using the same year data as an average

1. Get the data from the same year
There is a problem, after the joining, the original outliers were also included in the dataset. We need to **exclude** them.

```{r}
outliers_moving_average <- outliers |>
  select(-month, -value) |>
  left_join(d_hf_14_21 |>
  pivot_longer(cols = variable_columns, names_to = "index"))
```

2. exclude them use anti join
```{r}
fixed_outliers <- outliers_moving_average |> anti_join(outliers) |>
  group_by(state,
           stateuid,
           lga,
           lgauid,
           facility_name,
           facilityuid,
           year,
           index) |>
  summarise(value = mean(value, na.rm = TRUE))

fixed_outliers <-
  fixed_outliers |> left_join(outliers |> select(-value))

fixed_outliers |> write_csv(fixed_outliers_output)
```

3. Merging them back to database

```{r}
by_hf_year_month <- c(
  "state",
  "stateuid",
  "lga",
  "lgauid",
  "facility_name",
  "facilityuid",
  "year",
  "month",
  "index"
)
fixed_outliers <- read_csv(fixed_outliers_output)

d_hf_14_21 <- d_hf_14_21 |>
  pivot_longer(cols = variable_columns, names_to = "index") |>
  rows_update(fixed_outliers, by = by_hf_year_month) |>
  pivot_wider(names_from = "index", values_from = value) |>
  write_csv(file_output_fixed_outlier)
```

```{r}
d_hf_14_21 <- read_csv(file_output_fixed_outlier)
outliers_after_fixed <- find_outlier(
  d_hf_14_21,
  columns = variable_columns,
  alpha = 0.99999,
  select_rows = c(
          "state",
           "stateuid",
           "lga",
           "lgauid",
           "facility_name",
           "facilityuid",
           "year",
           "month"
           )
)
```

### TODO 3.0.2 Missing rate by indexes

Choose indexes

```{r}
missing_rate_by_indexes <- routine_monthly |>
  pivot_longer(
    cols = c(
      "test",
      "susp",
      "test_rdt",
      "test_mic",
      "test_rdt_lab",
      "abn_mic",
      "abn_rdt",
      "conf_rdt",
      "maltreat_u5",
      "maltreat_ov5",
      "maltreat",
      "maldth",
    ),
    names_to = "index",
    values_to = "value"
  ) |>
  mutate(reported = if_else(value > 0, 1, 0)) |>
  group_by(index) |>
  summarise(reports = sum(reported, na.rm = TRUE), n = n()) |>
  mutate(report_rate = 1 - reports / n)

ggplot(missing_rate_by_indexes) +
  geom_col(aes(y = fct_reorder(index, report_rate), x = report_rate), orientation =
             "y") +
  labs(title = "% missing indexes",
       y = "Indexes",
       x = "Percentage")
```

### TODO 3.0.3 Consistency test

#### 3.0.3.1 Suspect vs Test

```{r}
consis_susp_test <-
  routine_monthly |>
  mutate(susp_gt_test = if_else(susp < test, TRUE, FALSE)) |>
  ggplot() +
  geom_point(aes(y = susp, x = test, color = susp_gt_test)) +
  geom_abline(slope = 1)

print(consis_susp_test)
```

#### 3.0.3.2 test_rdt vs conf_rdt

```{r}
# change color
# and the legend
consis_test_conf_rdt <-
  routine_monthly |>
  mutate(test_rdt_lt_conf_rdt = if_else(test_rdt < conf_rdt, TRUE, FALSE)) |>
  ggplot() +
  geom_point(aes(y = test_rdt, x = conf_rdt, color = test_rdt_lt_conf_rdt)) +
  geom_abline(slope = 1)

print(consis_test_conf_rdt)
```

#### 3.0.3.3 conf vs maltreat

```{r}
consis_conf_maltreat <-
  routine_monthly |>
  mutate(conf_lt_maltreat = if_else(conf_rdt < maltreat, TRUE, FALSE)) |>
  ggplot() +
  geom_point(aes(y = conf_rdt, x = maltreat, color = conf_lt_maltreat)) +
  geom_abline(slope = 1)

print(consis_conf_maltreat)
```
