---
title: "Nigeria 2022 SNT"
editor_options:
  markdown:
    wrap: 72
output: pdf_document
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  comment = '', fig.width = 6, fig.height = 6,
  tidy = 'formatR',
  cache = TRUE
)
```

```{r include=FALSE}
devtools::load_all()
library(snt)
library(tidyverse)
library(haven)
library(readxl)
library(skimr)
```

```{r}
snt_country <<- "NGA"
```

# 1. cleaning

## 1.0 meta

### 1.0.1 input

```{r}
path_data <- "/Users/sepmein/Library/CloudStorage/OneDrive-SharedLibraries-WorldHealthOrganization/GMP-SIR\ -\ Country_Analytical_Support/Countries/NGA/WHO_NGA/NGA_2022_SNT/_Submitted_data/"
```

#### 1.0.1.1 Routine

```{r}
path_data_routine <- file.path(path_data, "Routine\ data")
path_data_routine_hf_1420 <- file.path(
    path_data_routine,
    "NGA_hflevel_2014-2020_raw.dta"
)
path_data_routine_hf_21 <- file.path(
    path_data_routine,
    "monthly_hf_2021.xlsx"
)
path_data_raw <- file.path(
  path_data_routine,
  "NGA_routineintdata_hf_month.dta"
)
```

### 1.0.2 output

```{r}
path_output_data <- "/Users/sepmein/Library/CloudStorage/OneDrive-SharedLibraries-WorldHealthOrganization/GMP-SIR\ -\ Country_Analytical_Support/Countries/NGA/WHO_NGA/NGA_2022_SNT/Final\ database\ for\ analysis"
```

## 1.1 merging before 2021 and the 2021 data

### 1.1.1 read health facility monthly data from 2014 - 2021

```{r}
d_hf_14_20 <- read_dta(path_data_routine_hf_1420)
```

#### Get colnames and skim the data

```{r}
colnames(d_hf_14_20)
skimr::skim(d_hf_14_20)
```

#### Show first rows

```{r}
head(d_hf_14_20)
```

### 1.1.2 read health facility monthly data of the year 2021

```{r echo = FALSE}
d_hf_21 <- read_xlsx(path_data_routine_hf_21)
```

#### Get colnames and skim the data

```{r}
colnames(d_hf_21)
skimr::skim(d_hf_21)
```

#### Show first N rows

```{r}
head(d_hf_21)
```

### 1.1.3 Data Ibrahim submitted on 11/06

```{r}
d_hf_21_raw <- read_dta(path_data_raw)
```

```{r}
colnames(d_hf_21_raw)
```

Merge with d_hf_21

```{r}
d_hf_21 |> left_join(d_hf_21_raw)
```

### 1.1.3 Reconstruct hf 2021

#### Rename the columns

```{r}
d_hf_21 <- d_hf_21 |>
  select(
    -fever_u5,
    -fever_a5,
    -tested_rdt_u5,
    -tested_rdt_a5,
    -rdtpos_u5,
    -rdtpos_a5,
    -conf_a5,
    -clinical_u5,
    -clinical_a5,
    -severe_u5,
    -severe_a5,
  ) |>
  rename(
    opd = allout,
    fever_u5  = susp_u5,
    fever_a5  = susp_ov5,
    tested_rdt_u5  = test_rdt_u5,
    tested_rdt_a5  = test_rdt_ov5,
    rdtpos_u5  = conf_rdt_u5,
    rdtpos_a5  = conf_rdt_ov5,
    tested_micro_a5  = test_mic_ov5,
    tested_micro_u5  = test_mic_u5,
    micropos_u5  = conf_mic_u5,
    micropos_a5  = conf_mic_ov5,
    #  clinical_preg  = pres_preg,
    clinical_u5  = pres_u5,
    clinical_a5  = pres_ov5,
    conf_a5  = conf_ov5,
    severe_a5  = maladm_ov5,
    severe_u5  = maladm_u5,
    ipd  = alladm,
    mal_death_u5  = maldth_u5,
    conf_u5 = "conf_u5...61",
    ipt1 = iptp1,
    ipt2 = iptp2,
    ipt3 = iptp3,
    ipt4 = iptp4,
    preg_llin = llins_anc
  )
```

#### Reorder the columns

```{r}
d_hf_21 <- d_hf_21 |>
  select(
    "facilityuid",
    "periodid",
    "periodcode",
    "perioddesc",
    "attendance",
    "opd",
    "fever_u5",
    "fever_a5",
    "tested_rdt_u5",
    "tested_rdt_a5",
    "rdtpos_u5",
    "rdtpos_a5",
    "tested_micro_u5",
    "tested_micro_a5",
    "micropos_u5",
    "micropos_a5",
    #"clinical_preg", # wait for Ibrahim to share data
    #"conf_preg", # wait for Ibrahim to share data
    "clinical_u5",
    "clinical_a5",
    "conf_u5",
    "conf_a5",
    "severe_u5",
    "severe_a5",
    "conf_act_u5",
    "conf_act_a5",
    "clinical_act_u5",
    "clinical_act_a5",
    "conf_othantimal_u5",
    "conf_othantimal_a5",
    "ipd",
    "mal_death_u5",
    #"stockout_act", # no need
    #"stockout_rdt", # no need
    #"stockout_sp", # no need
    #"stockout_llin", # no need
    "anc_total",
    "anc_b20wks",
    "anc_a20wks",
    "ipt1", 
    "ipt2", 
    "preg_llin",
    "fi_1yr_fixed",
    "fi_1yr_outreach",
    #"u5llin", 
    #"mal_ro", 
    #"mal_adr", 
    #stateuid, 
    #lgauid, 
    facility_name,
    "ownership",
    #"facilityname",
    #"longid",
    "year",
    "month",
    "state",
    "lga",
    "ward",
    "periodname"
  )
```

#### Fix the *periodname* column problem

When combining rows of two database: Error in `bind_rows()`: ! Can't
combine `..1$periodname` <date> and `..2$periodname` <character>. fix
that firstly

I found that the date information was already well stored in the
`periodcode` column, so what I am going to do is to **drop**
`periodname` column for both database.

```{r}
d_hf_14_20 <- d_hf_14_20 |> select( -periodname)
d_hf_21 <- d_hf_21 |> select(-periodname)
```

### 1.1.3 Combine Rows for the two dbs

bind rows between the old and new data

```{r}
d_hf_14_21 <- bind_rows(d_hf_14_20, d_hf_21)
```

### 1.1.4 Treat blank as NA

```{r}
d_hf_14_21 <- d_hf_14_21 |> mutate_all(na_if,"")

```

## 1.2 Cleaning

### 1.2.1 state and stateuid

As we can see from the merged data

```{r}
state_db <- d_hf_14_21 |>
  distinct(
    state,
    stateuid
  ) |>
  arrange(state)
state_db
```

We need to clean it a little bit. 1. Get a standard state name and uid
2. replace\_ it with the incorrect names 3. assign correct stateuid to
the according column

```{r}
# get a standard state name and uid

# remove unneccesary lines
state_db <- state_db |> 
  filter(!is.na(stateuid)) |> 
  filter(!nchar(stateuid) == 0)
  # use snt::routine_replace function to replace_ the state name
state_db <- snt::import_replace(
  state_db,
  snt::import_routine_adm1_replace,
  snt_country = snt_country,
  column = "state"
) |>
  distinct()
state_db
```

Now only have 37 rows.

Check the current \# of states in the db.

```{r}
d_hf_14_21 |> select(state) |> distinct() |> arrange(state)
```

Next, replace\_ all the state names in the database,

```{r}
d_hf_14_21 <- snt::import_replace(d_hf_14_21,
                                  snt::import_routine_adm1_replace,
                                  "NGA",
                                  "state") 

d_hf_14_21 |>
  select(state) |>
  distinct() |>
  arrange()
```

Remove the stateuid column then merge back the stata_db.

```{r}
d_hf_14_21 <- d_hf_14_21 |> 
  select(-stateuid) |>
  left_join(state_db)

d_hf_14_21 |> skim(stateuid)
```

Now 100% complete for stateuid. Fixed.

### 1.2.2 LGA names

Check the \# LGAs firstly.

```{r}
d_hf_14_21 |> select(lga) |> distinct() |> arrange(lga)
```

There are 1560 different LGAs for the original data.

The naming system in NGA for LGAs is like: "ak Mbo Local Government
Area", this should be changed to "Mbo"

```{r}
d_hf_14_21 <- d_hf_14_21 |>
  mutate_at(c("lga"), ~str_remove(., "^([a-z]{2} )")) |>
  mutate_at(c("lga"), ~str_remove(., " Local Government Area")) 
```

Check if it's better now:

```{r}
d_hf_14_21 |> distinct( lga ) |> arrange(lga)
```

I have created a database inside `snt` package for renaming, along with
a very useful function called `snt::import_replace` for easy replacing.

Standard `LGA/adm1` replacement from `snt` package looks like:

```{r}
nga_import_routine_adm2_replace <- snt::import_routine_adm2_replace |>
  filter(country == snt_country) |> arrange(from)
nga_import_routine_adm2_replace
```

Next, use the `snt::import_routine_adm2_replace` database to replace the
LGA column.

```{r}
# get routine rename database from snt database
d_hf_14_21 <- snt::import_replace(
  d_hf_14_21,
  snt::import_routine_adm2_replace,
  snt_country = snt_country,
  column = "lga"
)
```

Check if it's better now:

```{r}
d_hf_14_21 |> distinct( lga ) |> arrange(lga)
```

Next step, we need to rename the LGAs to align with those in shapefile.
The problem for NGA is sometimes, different adm1(state) has same LGA
name. E.g. both *Kogi* and *Plateau* has a LGA called *Bassa*. They
should be renamed with a number. country \| adm1 \| from \| to NGA \|
Kogi \| Bassa \| Bassa1\
NGA \| Plateau \| Bassa \| Bassa2

I have also recreated a db called
`snt::import_routine_adm2_replace_by_adm1`

replace LGAs to align the names in shapefile

```{r}
snt::import_routine_adm2_replace_by_adm1
```

Use `snt::import_replace` fn to replace those names.

```{r}
d_hf_14_21 <- snt::import_replace(
  d_hf_14_21,
  snt::import_routine_adm2_replace_by_adm1,
  snt_country = snt_country,
  column = "lga",
  by = c("state" = "adm1")
)
```

We can see the \# of lgas increased.

```{r}
d_hf_14_21 |> distinct(lga) |> arrange(lga)
```

#### TODO LGA UID

As 21 data do not have LGAs, I need to fix the problem. Skim over LGA
and lgauid column firstly.

```{r}
d_hf_14_21 |> skim(lga, lgauid)
```

Number of missing `lgauid` is 1677208

Print out \# of unique `lgauid` and `lga`.

```{r}
d_hf_14_21 |> distinct(lga, lgauid) |> arrange(lga)
```

As lots of lgauid are missing, filter those blank lines.

```{r}
lga_db <- d_hf_14_21 |> distinct(lga, lgauid) |> 
  arrange(lga) |>
  filter(!is.na(lgauid)) |> 
  filter(!nchar(lgauid) == 0)
lga_db
```

Now 774 rows.

Remove the `lgauid` column in `d_hf_14_21` db, and merge it back with
`lga_db`

```{r}
d_hf_14_21 <- d_hf_14_21 |> select(-lgauid) |> left_join(lga_db) 
d_hf_14_21 |>
  distinct(lga, lgauid) |> 
  arrange(lga) 
```

No missing LGAs, but found a few blank LGA

Print out those lines.

```{r}
d_hf_14_21 |> select(-lgauid) |> left_join(lga_db) |>
  filter(is.na(lga))
```

Try to find where they belongs to Search for 14-20 db

```{r}
facility_db <- d_hf_14_21 |>
  select(facilityuid,
         lga,
         lgauid) |>
  distinct()
facility_db
```

```{r}
d_hf_14_21 |>  
  select(
         -lga,
         -lgauid) |>
  left_join(facility_db) |>
  select(facilityuid,
         lga,
         lgauid) |>
  skim()
```

Still 47676 missing, failed

### 1.2.3 HF names

TODO later for RIA

### 1.2.4 Periods

Recreate the dropped `periodname` based on the `periodcode` column

```{r}
d_hf_14_21 |> skim()
```

# 2.0 Export

Export to the right position.

```{r}
file_output <- file.path(
  path_output_data,
  "hflevel_14-21.csv"
)
d_hf_14_21 |> write_csv(file_output)
```
